A Type System for Safe and Scalable Parallelism using Logical Regions

### 2012-10-26 doc/types-pldi2013/langdef.tex
checkpoint

@@

>>>>

\subsection{Serial Execution}
\label{subsec:coherenceserial}

@@

\subsection{Parallel Execution}
\label{subsec:coherenceparallel}

To determine when parallel execution is safe, we start from the serial execution trace and allow the reordering of 
adjacent heap operations that do not change the behavior of the application.  If we can show that reordering is allowed
between any two operations that come from different constituent traces, any interleaving of the constituent traces will
be equivalent to the serialized execution and parallel execution was safe.  Finally, to allow the safety of parallel
execution to be determined {\em before} the execution is actually performed, we show that a test based on the subtask
privileges and the current region mapping can soundly predict when this property will hold.

We define a {\em non-interference} operator $\epsilon_1 \nonint \epsilon_2$ on two memory operations as follows:
$$
\begin{array}{l@{ }l}
op_1(l_1, c_1,& v_1)\ \nonint\ op_2(l_2, c_2, v_2) \Leftrightarrow l_1 \not= l_2 \vee \\
& (op_1 = read \wedge op_2 = read)\ \vee \\
& (op_1 = reduce_{id_1} \wedge op_2 = reduce_{id_2} \wedge id_1 = id_2)
\end{array}
$$
Accesses to different locations cannot effect each other, and reads have no side-effects.  The safety of the third case
follows from the requirement that reduction operations be commutative.

The next step is to show that this non-interference operator is strong enough to preserve the validity of an interleaving
of two or more traces when reordering a pair of memory operations.
\begin{lem}
\label{lem:nonintswap}
\rm
Let $S$ be a store, $C$ a clobber set, $\oton{E}, E'_a, E'_b$ memory traces, and
$\epsilon_1, \epsilon_2$ be two memory operations from $E_i$ and $E_j$ ($i \not= j$).  Then,
$$\begin{array}{l@{}l}
valid\_interleave&(S, C, E'_a \dplus [ \epsilon_1, \epsilon_2 ] \dplus E'_b, \oton{E}) \wedge \epsilon_1 \nonint \epsilon_2 \Rightarrow \\
 & valid\_interleave(S, C, E'_a \dplus [ \epsilon_2, \epsilon_1 ] \dplus E'_b, \oton{E}).
\end{array}$$
\end{lem}
The proof follows from the definitions of {\em coherent} and {\em seq\_equiv} in Figure~\ref{fig:validinterleave}.

We now extend the notion of non-interference to whole memory traces, with two traces being non-interfering if no operation from one
trace interferes with any operation from the other:
$$E_1 \nonint E_2 \Leftrightarrow \displaystyle\bigwedge_{\epsilon_1 \text{ in } E_1, \epsilon_2 \text{ in } E_2} \epsilon_1 \nonint \epsilon_2$$

If whole memory traces are non-interfering, then any interleaving of them can be sorted via pairwise swaps to match the serial memory
trace and we get the result that shows parallel execution is safe:

\begin{lem}
\label{lem:noninteffects}
\rm
Let $S$ be an initial store, $C$ be a clobber set, $\oton{E}$ be memory traces such that $E_i \nonint E_j$
for every $1 \leq i < j \leq n$.  Then,
$any\_interleave(E', \oton{E}) \Rightarrow$ \\
\hspace*{0pt} \hfill $valid\_interleave(S, C, E', \oton{E})$.
\end{lem}

The final piece is to turn this test into something that can be performed before the subtasks are executed, ideally as efficiently as
possible.  The key to this is the privileges on Legion tasks and their ability to provide bounds on the accesses that may be performed
during execution.  Our non-interference operator is extended to work on privileges and sets of privileges:
$
\begin{array}{l@{}l}
\multicolumn{2}{l}{priv_1(r_1) \nonint[M] priv_2(r_2) \Leftrightarrow} \\
& (priv_1 = reads \wedge priv_2 = reads)\ \vee \\
& (priv_1 = reduces_{id_1} \wedge priv_2 = reduces_{id_2} \wedge id_1 = id_2)\ \vee \\
& (r_1 * r_2)\ \vee \\
& (M(r_1) \cap M(r_2) = \emptyset) \\
\\
\multicolumn{2}{l}{\Phi_1 \nonint[M] \Phi_2 \Leftrightarrow \displaystyle\bigwedge_{\phi_1 \in \Phi_1, \phi_2 \in \Phi_2} \phi_1 \nonint[M] \phi_2}
\end{array}$
Note although the
